AI Safety for High Energy Physics

### 2019-08-15 documentation/manuscript.tex
added text for the adversarial network methods

 \DeclareMathOperator{\sign}{sign}

@@

\newcommand{\epsilsonpt}{0.99}
\newcommand{\epsiloneta}{0.99}
\newcommand{\epsilonphi}{0.99}

@@

@@

\subsection{Fast Gradient Sign Method}
The FGSM method~\cite{DBLP:journals/corr/GoodfellowSS14} works by taking an individual jet $J_i$ and regressing the loss function to compute a \textit{bounded} perturbation $\delta J_i$, such that network's response to the input $J_i + \delta J_i$ tends towards some desired value.
Specifically, the perturbation is given by:
\begin{equation}
\label{eq:FGSM}
\delta J_i = \epsilon\left.\sign\left[\nabla_{J} \mathcal{L}\left(f(J), y_\mathrm{bg}\right)\right]\right\rvert_{J=J_i} \,,
\end{equation}
where $\mathcal{L}$ is the binary crossentropy, and $y_\mathrm{bg}$ is the label corresponding to background events.
The use of the $\sign$ of the gradient and the free parameter $\epsilon$ ensures that the perturbation of each input observable is bounded while moving approximately in the direction normal to the decision boundary.

@@

A deep neural network $f(J_i)\mapsto [0,1]$ is trained in the usual way $f=\text{argmin}_{f'}\mathcal{L}(F'(J),y)$, $y\in\{0,1\}$, where $\mathcal{L}$ is the loss function (in this case, cross-entropy).
Networks are trained with Keras~\cite{keras} using the Tensorflow~\cite{tensorflow} backend and optimized using Adam~\cite{adam}.
To compare the sensitivity of a deep neural network trained directly on the natural high-dimensionality with traditional methods, a second network is trained whose first layer is a ratio of energy correlation functions~\cite{Larkoski:2013eya} known as $D_2$~\cite{Larkoski:2014gra}.

>>>>

 \DeclareMathOperator{\sign}{sign}
 \DeclareMathOperator*{\argmin}{argmin}
 \DeclareMathOperator*{\E}{\mathbb{E}}

@@

\newcommand{\epsilsonpt}{0.99} % Epsilon values for pt/eta/phi in FGSM attack
\newcommand{\epsiloneta}{0.99}
\newcommand{\epsilonphi}{0.99}

%adversarial network parameters
\newcommand{\nlayerADV}{3}
\newcommand{\nunitsADV}{512}

@@

\todo{rewrite to include information about second benchmark classifier that takes in jet-level observables (pT, eta, phi, mass, D2)}

@@

\subsection{Fast Gradient Sign Method}
\label{subsec:fgsm}
The FGSM method~\cite{DBLP:journals/corr/GoodfellowSS14} works by taking an individual jet $J_i$ and regressing the loss function to compute a \textit{bounded} perturbation $\delta J_i$, such that network's response to the input $J_i + \delta J_i$ tends towards some desired value.
Specifically, the perturbation is given by:
\begin{equation}
\label{eq:FGSM}
\delta J_i = \epsilon\left.\sign\left[\nabla_{J} \mathcal{L}\left(f(J), y_\mathrm{bg}\right)\right]\right\rvert_{J=J_i} \,,
\end{equation}
where $\mathcal{L}$ is the binary crossentropy, and $y_\mathrm{bg}$ is the label corresponding to background events.
The use of the $\sign$ of the gradient and the free parameter $\epsilon$ ensures that the perturbation of each input observable is bounded while moving approximately in the direction normal to the decision boundary.

@@

In this approach, we train a second neural network $g$ for which both input and output are $\nconst\times 3$ tensors.
The goal is to train $g$ to map $J\mapsto\delta J$, for an arbitrary jet such that the classifier network $f$ will perceive the input $J+\delta J$ as a background jet.
The network $g$ is trained by minimizing the loss function:
\begin{multline}
g = \argmin_{g'}\E_{J\sim data} [ \mathcal{L}_{xe}(f(J+g'(J)), y_\mathrm{bg}) \\ + \lambda\mathcal{L}_{obs}(J, g'(J))] \,.
\end{multline}
Here, $\mathcal{L}_{xe}$ is the usual binary crossentropy loss function, and $\mathcal{L}_{obs}$ is in our example a family of loss functions that minimize the difference of jet-level observables between $J$ and $J+\delta J$.
$\mathcal{L}_{obs}$ may also contain terms involving only $g'(J)$, in particular, it may include $L_2$ activity regularizers that encourage the perturbations output by network $g$ to be small.
The parameter $\lambda$ controls how aggressively the adversary $g$ will attempt to confound the classifier $f$, versus preserving the jet-level observables.

In our experiments, $g$ is a fully-connected network with \nlayerADV\ hidden layers, each with \nunitsADV\ units and ReLU activation.
The final output layer has $\nconst\times 3$ units, with $\tanh$ activations.
Analogously to the bounding parameters $\epsilon$ in \ref{subsec:fgsm}, the outputs of the final layer are bounded by applying a $\tanh$ activation functions and the axes corresponding to $\pt$, $\eta$, and $\phi$ are scaled by parameters $\rho_{\pt}$, $\rho_\eta$, and $\rho_\phi$ respectively.

%To compare the sensitivity of a deep neural network trained directly on the natural high-dimensionality with traditional methods, a second network is trained whose first layer is a ratio of energy correlation functions~\cite{Larkoski:2013eya} known as $D_2$~\cite{Larkoski:2014gra}.
